---
title: "p8105_hw6_tl3195"
author: "Tingyi Li"
date: "2023-12-01"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(mgcv)
library(modelr)
set.seed(8105)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 2

*Import the dataset 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

*A simple linear regression with tmax as the response with tmin and prcp as the predictors
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

boot_straps = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df)),
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data=df)),
    r_squared = map(models, broom::glance),
    results = map(models, broom::tidy)
  )|>
  select(strap_number, r_squared, results)|>
  unnest(r_squared, results)|>
  select(strap_number, r.squared, term, estimate)|>
  pivot_wider(
    names_from = term,
    values_from = estimate
  )|>
  janitor::clean_names()|>
  mutate(
    logb = log(prcp * tmin)
  )|>
  select(strap_number, r_squared, logb)

```

*Plot the distribution of r_squared
```{r}
boot_straps|>
  ggplot(aes(x = r_squared)) +
  geom_density()
```

*Plot the distribution of logb
```{r}
boot_straps|>
  filter(!is.na(logb))|>
  ggplot(aes(x = logb)) +
  geom_density()
```

The first plot shows the distribution of the r-squared values from the data, it peaks at around 0.91. The second plot shows the distribution of log(b1*b2) from the data, it peaks at around -5.5. Both the plots are left-skewed but the first plot appears slightly skewed to the left. 


*95% CI for r_squared and logb
```{r}
boot_straps|>
  filter(!is.na(logb))|>
  pivot_longer(
    r_squared:logb,
    names_to = "term",
    values_to = "estimate"
  )|>
  group_by(term)|>
  summarize(
    ci_lower = quantile(estimate, 0.025),
    ci_upper = quantile(estimate, 0.975)
  )|>
  knitr::kable(digits = 3)
```


#Problem 3

*Load and clean the data for regression analysis
```{r}
birth_weight = read_csv("./data/birthweight.csv")|>
  janitor::clean_names()|>
  mutate(
    babysex = as.factor(babysex),
    babysex = recode(babysex, `1` = "male", `2` = "female"),
    frace = as.factor(frace),
    frace = recode(frace, `1` = "White", `2` = "Black", `3` = "Asian", `4` = "Puerto Rican", 
                   `8` = "Other", `9` = "Unknown"),
    mrace = as.factor(mrace),
    mrace = recode(mrace, `1` = "White", `2` = "Black", `3` = "Asian", `4` = "Puerto Rican", 
                   `8` = "Other"),
    malform = as.factor(malform),
    malform = recode(malform, `0` = "absent", `1` = "present"),
  )
```

*Selecting Variables 
```{r}
full_model = lm(bwt  ~., data = birth_weight)

step(full_model, direction = 'backward', trace = FALSE)
```

When selecting variables, I use the backward selection method, which start with the full model and remove the least significant variables.

Therefore, the final model should be:
```{r}
final_model = lm(formula = bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + mheight + mrace + parity + ppwt + smoken, data = birth_weight)

birth_weight|>
  add_predictions(final_model)|>
  add_residuals(final_model)|>
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  labs(
    title = "Residuals vs Fitted values",
    x = "Fitted values",
    y = "Residuals"
  )
```

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
cv_df = 
  crossv_mc(birth_weight, 100) |>
  mutate(
    maineffect_model = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    complex_model =  map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x)),
    final_model = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + 
                         mheight + mrace + parity + ppwt + smoken, data = .x))
    )|>
  mutate(
    rmse_maineffect = map2_dbl(maineffect_model, test, rmse),
    rmse_complex = map2_dbl(complex_model, test, rmse),
    rmse_final = map2_dbl(final_model, test, rmse)
    ) 
```

```{r}
cv_df|>
  select(starts_with("rmse"))|>
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_")|>
  mutate(model = fct_inorder(model))|>
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

